2H qSIP Calculations

### 2021-11-15 qSIP_calculations.Rmd
removed practice plots moved to figure script

gen_calc_data_summarized <- left_join(gen_calc_data_summarized, 
                                     weighted_turnover_summarized,
                                     by = c("terrain", "inc_time_d"))

# gen_calc_data_summarized %>% openxlsx::write.xlsx(file.path("data", "gen_calc_data_summarized.xlsx"))
```

@@

# Practice Plotting

## Peak Tracer Incorporation
```{r}
peaks_mapped_with_zeros_analytes %>% ggplot(
  aes(x = inc_time_d, y = at2H, color = compound, group = compound)) +
  geom_point() +
  geom_smooth(se = FALSE,
              method = "lm") +
  #geom_borderline() +
  scale_color_viridis_d() +
  theme_minimal()
```

## Chromatogram scaled by 2F

```{r "shaded chromatogram", fig.width=8, fig.height=8}
p_peaks_mapped_active <- peaks_mapped_with_zeros_analytes %>% ggplot(
  aes(x = rt, y = amp2, fill = as.numeric(at2H), label = compound)) +
  geom_col(aes(width = rt_end - rt_start)) +
  scale_fill_viridis_c(option = "magma") +
  facet_wrap(~sample_id, ncol = 1) +
  geom_text_repel(size = 2, box.padding = 0.5, nudge_y = 150) +
  theme_classic() +
  coord_cartesian(ylim = c(0, 1300)) +
  labs(fill = TeX("$^2F$"),
         x = "Retention Time (s)",
         y = "Amplitude (mV)") +
  theme(strip.background = element_blank())

## Compound enrichment over time

```{r}
peaks_mapped_summarized %>% 
  mutate(compound = forcats::fct_reorder(compound, rt)) %>% 
  ggplot(aes(x = inc_time_d,
             y = at2H,
             group = compound)) +
  geom_borderline(color = "gray") +
  geom_point() +
  geom_line(aes(x = inc_time_d,
                y = at2H_weighted,
                color = terrain),
            size = .8) +
  geom_point(aes(x = inc_time_d,
                y = at2H_weighted, 
                color = terrain),
             shape = 18,
             size = 3) +
  geom_label_repel(data = peaks_mapped_summarized %>% filter(inc_time_d == 7),
                   nudge_x = 2,
                   nudge_y = 0.005,
                   size = 2.5,
                   aes(label = compound)) +
  geom_label_repel(data = weighted_time_course %>% filter(inc_time_d == 3),
                   aes(x = inc_time_d,
                       y = at2H_weighted,
                       group = terrain,
                       color = terrain),
                   label = "Weighted Mean",
                   nudge_y = -.01,
                   nudge_x = 1,
                   size = 2.5) +
  facet_wrap(vars(terrain), nrow = 1) +
  scale_x_continuous(breaks = c(0, 3, 7)) +
  scale_color_npg() +
  coord_cartesian(xlim = c(0, 10), clip = FALSE) +
  labs(x = "Incubation Time (Days)",
       y = TeX("% $^2H$"),
       color = "Soil Type") +
  theme_classic() +
  theme(legend.position = "none",
        strip.background = element_blank())

## Compound enrichment over time with label strength

```{r}
gen_calc_data_summarized %>% 
  filter(compound != "i-16:0") %>% 
  mutate(compound = fct_reorder(compound, rt)) %>% 
  ggplot(aes(x = inc_time_d,
             y = at2H,
             color = compound)) +
  geom_point() +
  geom_borderline() +
  geom_hline(yintercept = 0.5,
             color = "red") +
  facet_wrap(vars(terrain)) +
  scale_x_continuous(breaks = c(0, 3, 7)) +
  scale_color_viridis_d() +
  labs(x = "Incubation Time (Days)",
       y = "Mean %D") +
  theme_bw()
```

## Generation Time of All Points

```{r}
p_gen_time_dist <- gen_calc_data %>% 
  filter(inc_time_d != 0) %>% 
  ggplot(
  aes(y = gen_d, 
      x = inc_time_d_str, 
      color = inc_time_d_str, 
      fill = inc_time_d_str)) +
  stat_slab(width = 0.3, alpha = 0.7) +
  stat_pointinterval(color = "black") +
  # geom_boxplot(fill = "white",
  #              color = "black",
  #              width = 0.1,
  #              outlier.shape = NA) +
  geom_point(
    color = "black",
    shape = 95,
    size = 3,
    alpha = 0.5,
    position = position_nudge(x = -.03)) +
  # geom_text_repel(aes(label = compound), 
  #                 color = "black", 
  #                 nudge_y = -0.3, 
  #                 size = 3) +
  scale_y_log10() +
  annotation_logticks(sides = "l") +
  scale_fill_npg() +
  scale_color_npg() +
  theme_classic() +
  labs(y = "Generation time (days)",
       x = "Incubation time",
       title = "Generation times of soil bacteria",
       subtitle = "Lipid-specific generation times of bacteria",
       caption = "Soils collected from Colorado") +
  theme(legend.position = "none",
        axis.ticks.y = element_blank())
p_gen_time_dist

#ggsave("fig_output/p_gen_times_dist.png", plot = p_gen_time_dist, height = 6, width = 8, dpi = 300)
```

## Generation Time, all points, not separated by inc_time

```{r}
p_gen_time_dist_all <- gen_calc_data %>% filter(inc_time_d != 0) %>% 
  ggplot(
  aes(y = gen_d, 
      x = 1)) +
  stat_slab(width = 0.3, alpha = 0.7) +
  stat_pointinterval(color = "black") +
  # geom_boxplot(fill = "white",
  #              color = "black",
  #              width = 0.1,
  #              outlier.shape = NA) +
  geom_point(
    color = "black",
    shape = 95,
    size = 3,
    alpha = 0.5,
    position = position_nudge(x = -.03)) +
  # geom_text_repel(aes(label = compound), 
  #                 color = "black", 
  #                 nudge_y = -0.3, 
  #                 size = 3) +
  scale_y_log10() +
  annotation_logticks(sides = "l") +
  scale_fill_npg() +
  scale_color_npg() +
  theme_classic() +
  labs(y = "Generation time (days)",
       title = "Generation times of soil bacteria",
       subtitle = "Lipid-specific generation times of bacteria",
       caption = "Soils collected from Colorado") +
  theme(legend.position = "none",
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank())
p_gen_time_dist_all
```

## Generation Time, all points, not separated, > C19 removed

```{r}
p_gen_time_dist_all_no19 <- gen_calc_data %>% 
  filter(inc_time_d != 0,
         chain_num < 19) %>% 
  ggplot(
  aes(y = gen_d, 
      x = 1)) +
  stat_slab(width = 0.3, alpha = 0.7) +
  stat_pointinterval(color = "black") +
  # geom_boxplot(fill = "white",
  #              color = "black",
  #              width = 0.1,
  #              outlier.shape = NA) +
  geom_point(
    color = "black",
    shape = 95,
    size = 3,
    alpha = 0.5,
    position = position_nudge(x = -.03)) +
  # geom_text_repel(aes(label = compound), 
  #                 color = "black", 
  #                 nudge_y = -0.3, 
  #                 size = 3) +
  scale_y_log10() +
  annotation_logticks(sides = "l") +
  scale_fill_npg() +
  scale_color_npg() +
  theme_classic() +
  labs(y = "Generation time (days)",
       title = "Generation times of soil bacteria",
       subtitle = "Lipid-specific generation times of bacteria \nC20+ removed",
       caption = "Soils collected from Colorado") +
  theme(legend.position = "none",
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank())
p_gen_time_dist_all_no19
```

## Generation time of all points, not log scale

```{r}
p_gen_time_dist_all_lin <- gen_calc_data %>% 
  filter(inc_time_d != 0,
         gen_d >= 0,
         chain_num < 19) %>% 
  ggplot(
  aes(y = gen_d, 
      x = 1)) +
  stat_slab(width = 0.25, 
            alpha = 0.7,
            color = "black") +
  stat_pointinterval(color = "black") +
  # geom_boxplot(fill = "white",
  #              color = "black",
  #              width = 0.1,
  #              outlier.shape = NA) +
  geom_point(
    color = "black",
    shape = 95,
    size = 3,
    alpha = 0.5,
    position = position_nudge(x = -.03)) +
  # geom_text_repel(aes(label = compound), 
  #                 color = "black", 
  #                 nudge_y = -0.3, 
  #                 size = 3) +
  scale_fill_npg() +
  scale_color_npg() +
  theme_classic() +
  labs(y = "Generation time (days)",
       title = "Generation times of soil bacteria",
       subtitle = "Lipid-specific generation times of bacteria",
       caption = "Soils collected from Colorado") +
  theme(legend.position = "none",
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank(),
        axis.title.x = element_blank())
p_gen_time_dist_all_lin
```

## Generation Time, Broken by Terrain

```{r}
signif_comparisons <- gen_calc_data %>% pull(terrain) %>% unique()

p_gen_time_terrain <- gen_calc_data %>% 
  filter(inc_time_d != 0,
         chain_num < 19) %>% 
  ggplot(
  aes(y = gen_d, 
      x = terrain,
      color = terrain,
      fill = terrain)) +
  stat_slab(width = 0.25, 
            alpha = 0.7,
            color = "black") +
  stat_pointinterval(color = "black") +
  # geom_boxplot(fill = "white",
  #              color = "black",
  #              width = 0.1,
  #              outlier.shape = NA) +
  geom_point(
    color = "black",
    shape = 95,
    size = 3,
    alpha = 0.5,
    position = position_nudge(x = -.07)) +
  geom_signif(comparisons = list(signif_comparisons),
              map_signif_level = TRUE) +
  scale_y_log10(breaks = c(1, 10, 100)) +
  annotation_logticks(sides = "l") +
  scale_fill_npg() +
  scale_color_npg() +
  theme_classic() +
  labs(y = "Generation time (days)",
       title = "Generation times of soil bacteria",
       subtitle = "Lipid-specific generation times of bacteria \nC20+ omitted") +
  theme(axis.ticks.y = element_blank(),
        axis.text.x = element_text(angle = 45, hjust = 1))
p_gen_time_terrain
```

## Turnover time by compound

```{r}
gen_calc_data_summarized %>% 
  filter(inc_time_d != 0, compound != "i-16:0") %>%
  group_by(compound) %>% 
  summarize_if(is.numeric, mean, na.rm = TRUE) %>% 
  mutate(compound = fct_reorder(compound, u_d)) %>% 
  ggplot(aes(x = compound,
             y = u_d,
             fill = u_d)) +
  geom_col(position = "dodge") +
  scale_fill_viridis_c(option = "plasma") +
  labs(x = "Compound",
       y = "Mean Turnover time (d^-1)",
       title = "Compound-specific turnover rates of bacterial phospholipid fatty acids") +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, 
                               hjust = 1)
  )
```

## Turnover Time, Broken By Terrain

```{r}
p_turnover_terrain <- gen_calc_data %>% filter(inc_time_d != 0) %>% 
  ggplot(
  aes(x = inc_time_d_str, 
      y = u_d, 
      color = terrain, 
      fill = terrain)) +
  #stat_slab(height = 0.3, alpha = 0.7) +
  stat_pointinterval(color = "black") +
  # geom_point(
  #   color = "black",
  #   #shape = 124,
  #   size = 1,
  #   alpha = 0.5,
  #   position = position_nudge(x = -.05)) +
  # geom_text_repel(aes(label = compound), 
  #                 color = "black", 
  #                 nudge_y = -0.3, 
  #                 size = 3) +
  facet_wrap(~terrain, nrow = 1) +
  scale_y_log10() +
  annotation_logticks(sides = "a") +
  scale_fill_npg() +
  scale_color_npg() +
  theme_bw() +
  labs(x = "Incubation Time",
       y = "Turnover Time (d^-1)",
       title = "Generation times of soil bacteria across terrain")
  theme(legend.position = "none",
        axis.ticks.y = element_blank())
p_turnover_terrain
```

## 7 Day Turnover Times By Terrain

```{r}
p_turnover_terrain_7d <- gen_calc_data %>% filter(inc_time_d == 7) %>% 
  ggplot(
  aes(x = terrain, 
      y = u_d)) +
  #stat_slab(height = 0.3, alpha = 0.7) +
  #stat_pointinterval(color = "black") +
  geom_boxplot(outlier.shape = NA) +
  geom_point(
    color = "black",
    #shape = 124,
    size = 1,
    alpha = 0.5) +
  geom_signif(comparisons = list(signif_comparisons),
              map_signif_level = TRUE) +
  scale_y_log10() +
  annotation_logticks(sides = "a") +
  scale_fill_npg() +
  scale_color_npg() +
  theme_bw() +
  labs(x = "Terrain",
       y = "Turnover time (d^-1)",
       title = "Generation times of soil bacteria across terrain") +
  theme(legend.position = "none",
        axis.ticks.y = element_blank())
p_turnover_terrain_7d
```

## 7 Day Generation Times across terrain

```{r}
p_gen_time_terrain_7d <- gen_calc_data %>% filter(inc_time_d == 7) %>% 
  ggplot(
  aes(x = terrain, 
      y = gen_d)) +
  #stat_slab(height = 0.3, alpha = 0.7) +
  #stat_pointinterval(color = "black") +
  geom_boxplot(outlier.shape = NA) +
  geom_point(
    color = "black",
    #shape = 124,
    size = 1,
    alpha = 0.5) +
  # geom_text_repel(aes(label = compound), 
  #                 color = "black", 
  #                 nudge_y = -0.3, 
  #                 size = 3) +
  scale_y_log10() +
  annotation_logticks(sides = "a") +
  scale_fill_npg() +
  scale_color_npg() +
  theme_bw() +
  labs(x = "Terrain",
       y = "Turnover time (d^-1)",
       title = "Generation times of soil bacteria across terrain") +
  theme(legend.position = "none",
        axis.ticks.y = element_blank())
p_gen_time_terrain_7d
```

## Turnover time versus lipid relative abundance

gen_calc_data %>% 
  filter(inc_time_d != 0) %>% 
  ggplot(aes(x = rel_area_FID,
             y = u_d)) +
  geom_point(aes(color = chain_fctr)) +
  geom_smooth(method = "lm",
              se = FALSE) +
  scale_color_viridis_d() +
  facet_wrap(~terrain) +
  labs(x = "Relative Area (GC-FID)",
       y = "Turnover (d^-1)",
       color = "Compound class") +
  theme_bw()

## Compound-Specific Enrichment

```{r}
gen_calc_data %>% 
  mutate(compound = fct_reorder(compound, -chain_num)) %>% 
  ggplot(aes(x = at2H,
             y = compound,
             fill = stat(x))) +
  geom_density_ridges_gradient(scale = 3) +
  # Vline showing enrichment strength
  # geom_vline(xintercept = 0.5,
  #            color = "red") +
  facet_wrap(vars(inc_time_d_str), ncol = 1) +
  scale_fill_viridis_c(option = "C") +
  labs(fill = "% D",
       x = "% D",
       y = "Compound") +
  theme_ridges() +
  theme(axis.text.y = element_text(size = 5),
        legend.position = "right",
        legend.key.width = unit(0.2, "line"),
        legend.text = element_text(size = 10))
```

## Compound specific turnover
```{r}
gen_calc_data %>% 
  mutate(compound = fct_reorder(compound, -chain_num)) %>% 
  ggplot(aes(x = u_d,
             y = compound,
             fill = stat(x))) +
  geom_density_ridges_gradient() +
  # Vline showing enrichment strength
  facet_wrap(vars(inc_time_d_str), ncol = 1) +
  scale_fill_viridis_c(option = "C") +
  labs(fill = "µ (d^-1)",
       x = "µ (d^-1)",
       y = "Compound") +
  theme_minimal()
```

## Mean Turnover by Terrain

```{r}
gen_calc_data_summarized %>% 
  group_by(inc_time_d, terrain) %>% 
  summarize(u_d_mn = mean(u_d),
            u_d_max = max(u_d),
            u_d_min = min(u_d),
            gen_d_mn = mean(gen_d),
            gen_d_max = max(gen_d),
            gen_d_min = min(gen_d)) %>% 
  ggplot(aes(x = terrain, 
             y = u_d_mn,
             color = as.character(inc_time_d))) +
  geom_path(color = "black") +
  geom_point() +
  coord_cartesian(ylim = c(0, 0.05)) +
  labs(x = "Terrain",
       y = "Mean Turnover (d^-1)",
       color = "Incubation Time") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

## Mean Turnover Aggregated

```{r}
gen_calc_data_summarized %>% 
  # Calculate means, by terrain
  group_by(inc_time_d, terrain) %>% 
  summarize(u_d_mn = mean(u_d),
            u_d_max = max(u_d),
            u_d_min = min(u_d),
            gen_d_mn = mean(gen_d),
            gen_d_max = max(gen_d),
            gen_d_min = min(gen_d)) %>% 
  # Calculate mean of means
  group_by(inc_time_d) %>% 
  summarize(u_d_mn_ag = mean(u_d_mn),
            u_d_max_ag = max(u_d_mn),
            u_d_min_ag = min(u_d_mn),
            gen_d_mn_ag = mean(gen_d_mn),
            gen_d_max_ag = max(gen_d_mn),
            gen_d_min_ag = min(gen_d_mn)) %>% 
  ggplot(aes(x = 1, 
             y = u_d_mn_ag,
             color = as.character(inc_time_d))) +
  geom_path(color = "black") +
  geom_point() +
  coord_cartesian(ylim = c(0, 0.05)) +
  labs(x = "Terrain",
       y = "Mean Turnover (d^-1)",
       color = "Incubation Time") +
  theme_minimal() +
  theme(axis.text.x = element_blank(),
        axis.title.x = element_blank())

## Literature Rates of Cell Turnover

```{r}
# Import lit search data
lit_turnover <- read_xlsx(file.path("data", "literature_turnover_rates.xlsx"))

lit_turnover <- lit_turnover %>% 
  # Arrange by turnover time
  arrange(u_d_mn) %>% 
  # Reassign n values
  mutate(n = seq(from = 1, to = nrow(lit_turnover)))

# Our summarized data
gen_calc_mini <- gen_calc_data_summarized %>% 
  # Calculate mean per terrain per inc time
  group_by(inc_time_d, terrain) %>% 
  summarize(u_d_mn = mean(u_d, na.rm = TRUE),
            u_d_max = max(u_d),
            u_d_min = min(u_d)) %>%
  # Calculate mean of means (combine inc_times)
  group_by(terrain) %>% 
  summarize(u_d_mn = mean(u_d_mn, na.rm = TRUE),
            u_d_max = max(u_d_mn, na.rm = TRUE),
            u_d_min = min(u_d_mn, na.rm = TRUE)) %>% 
  # Rename to match `lit_turnover` tibble
  rename(soil_type = terrain) %>% 
  arrange(u_d_mn) %>% 
  mutate(method = "2H Lipid SIP",
         n = c(17, 18, 19),
    soil_type = case_when(
      soil_type == "Niwot Ridge Tundra" ~ "tundra",
      soil_type == "Gordon Gulch Conifer Forest" ~ "conifer forest",
      soil_type == "Marshall Mesa Grassland" ~ "grassland")
  )

# Add in our data! ^_^
lit_turnover_w_ours <- lit_turnover %>% bind_rows(gen_calc_mini)

lit_turnover_w_ours <- lit_turnover_w_ours %>% 
  mutate(
    source = case_when(
      method == "2H Lipid SIP" ~ "This study",
      TRUE ~ "Literature values"
  ))

# By terrain

lit_turnover_w_ours %>%
  # order by mean turnover rate
  ggplot(aes(x = n, 
             y = u_d_mn,
             color = soil_type)) +
  geom_segment(aes(y = 0, yend = u_d_mn, xend = n),
               color = "black") +
  geom_point(size = 3) +
  scale_color_npg() +
  geom_label(label = "This study",
            x = 18,
            y = 0.13,
            color = "black") +
  geom_label(label = "Literature values",
            x = 9,
            y = 0.6,
            color = "black") +
  #scale_y_log10() +
  #annotation_logticks(sides = "l") +
  coord_cartesian(
    xlim = c(0, 20),
    ylim = c(0, 0.80), 
    expand = FALSE) +
  labs(y = "Mean Turnover Time (d^-1)",
       color = "Soil Source") +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "bottom")

## With log scale

lit_turnover_w_ours %>%
  # order by mean turnover rate
  ggplot(aes(x = n, 
             y = u_d_mn,
             color = soil_type)) +
  geom_segment(aes(y = 0, yend = u_d_mn, xend = n),
               color = "black") +
  geom_point(size = 3) +
  scale_color_npg() +
  geom_label(label = "This study",
            x = 18,
            y = 0.10,
            color = "black") +
  geom_label(label = "Literature values",
            x = 9,
            y = 0.35,
            color = "black") +
  scale_y_log10() +
  annotation_logticks(sides = "l") +
  labs(y = "Mean Turnover Time (d^-1)",
       color = "Soil Source") +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        legend.position = "bottom")

## Literature Rates of Cell Turnover by Method
```{r}
lit_turnover_w_ours %>%
  # order by mean turnover rate
  ggplot(aes(x = n, 
             y = u_d_mn,
             color = method)) +
  geom_segment(aes(y = 0, yend = u_d_mn, xend = n),
               color = "black") +
  geom_point(size = 3) +
  geom_text_repel(aes(label = method),
                  color = "black",
                  size = 2,
                  segment.color = "red",
                  nudge_y = 0.2,
                  segment.curvature = -0.1,
                  segment.angle = -20,
                  segment.ncp = 3,
                  segment.size = 0.1) +
  scale_color_viridis_d() +
  scale_y_log10(breaks = c(0.01, 0.05, 0.1, 0.5)) +
  annotation_logticks(sides = "l") +
  labs(y = "Mean Turnover Time (d^-1)",
       color = "Soil Source") +
  theme_classic() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title.x = element_blank(),
        )
```

>>>>

```{r join weighted gen calc}
gen_calc_data_summarized_weighted <- left_join(gen_calc_data_summarized, 
                                     weighted_turnover_summarized,
                                     by = c("terrain", "inc_time_d"))

@@
